---
title: "CytB Differences"
output: html_notebook
---

Make a matrix of cytb differences.

I think I can do this with Ape and the Unique Sequences phylip file


```{r}
pacman::p_load(ape, tidyverse, kableExtra)
```


```{r}
dn <- read.dna('cytb.nex.uniqueseq.phy')
```

```{r}
m.dn <- round(dist.dna(dn, as.matrix = T, model = 'JC69', pairwise.deletion = T), 4) %>% as_tibble(rownames = NA) %>% 
  rownames_to_column("ID") %>% 
  write_csv("cytb_distances.csv")
```

I want to change all the names so that just the species is present.

```{r}
 # THis removes everything after first period
m.dn2 <- gsub("^.*?\\.","", m.dn$ID ) %>% 
  as_tibble(rownames = NA) %>% 
  rownames_to_column("ID")

#after second period
m.dn3 <- gsub("\\..*","",m.dn2$value) %>% 
  as_tibble(rownames = NA) %>% 
  mutate(value = recode(value, "cf" = "cf.neglecta"))

#m.dn.g <- 
  
difs <- m.dn %>% bind_cols(m.dn3) %>% 
  select(!ID) %>% 
  gather("sample", "dif", -value) %>% 
  mutate(sample = gsub("^.*?\\.","", sample )) %>% 
  mutate(sample = gsub("\\..*","", sample )) %>%
  mutate(sample = recode(sample, "cf" = "cf.neglecta")) %>% 
  filter(sample %in% c('balingka', 'barapi', 'dewi', 'neglecta', 'cf_neglecta', 'hutanis', 'lepidura', 'monticola', 'umbra', 'abscondita', 'paradoxura', 'beccarii', 'maxi')) %>% 
  filter(value %in% c('balingka', 'barapi', 'dewi', 'neglecta', 'cf_neglecta', 'hutanis', 'lepidura', 'monticola', 'umbra', 'abscondita', 'paradoxura', 'beccarii', 'maxi')) %>% 
  unite(combo, value, sample, sep = "_", remove = FALSE) %>% 
  select(!c(value, sample)) %>% 
  group_by(combo) %>% 
  summarise(mean = mean(dif),
            min = min(dif),
            max = max(dif)) %>% 
  mutate_if(is.numeric, ~round(., 2)) %>% 
  #mutate_at(c(2:3), ~replace(., is.na(.), 0)) %>% 
  unite(range, min, max, sep = ",", remove = FALSE) %>% 
  select(!c(min, max)) %>% 
  unite(mean, mean, range, sep = " (", remove = FALSE) %>% 
  select(!range) %>% 
  separate(combo, into = c('a', 'b'), sep = '_')


```

# Get mean differences
```{r}
difs2 <- m.dn %>% bind_cols(m.dn3) %>% 
  select(!ID) %>% 
  gather("sample", "dif", -value) %>% 
  mutate(sample = gsub("^.*?\\.","", sample )) %>% 
  mutate(sample = gsub("\\..*","", sample )) %>%
  mutate(sample = recode(sample, "cf" = "cf.neglecta")) %>% 
  filter(sample %in% c('balingka', 'barapi', 'dewi', 'neglecta', 'cf_neglecta', 'hutanis', 'lepidura', 'monticola', 'umbra', 'abscondita', 'paradoxura', 'beccarii', 'maxi')) %>% 
  filter(value %in% c('balingka', 'barapi', 'dewi', 'neglecta', 'cf_neglecta', 'hutanis', 'lepidura', 'monticola', 'umbra', 'abscondita', 'paradoxura', 'beccarii', 'maxi')) %>% 
  unite(combo, value, sample, sep = "_", remove = FALSE) %>% 
  select(!c(value, sample)) %>% 
  group_by(combo) %>% 
  summarise(mean = mean(dif)) %>% 
  mutate_if(is.numeric, ~round(., 2)) %>% 
  separate(combo, into = c('a', 'b'), sep = '_')

mat <- xtabs(mean ~ a + b, difs2) %>% matrixcalc::upper.triangle() %>% as.data.frame.matrix()

mat2 <- xtabs(mean ~ a + b, difs2)
mat2[upper.tri(mat2)] <- ""
mat2 <- mat2 %>% as.data.frame.matrix() %>% as_tibble(rownames = NA) %>% write.csv('distance_matrix.csv')
```

# Print matrix
I manually added the values and saved it as a csv called `distance_matrix_range.csv`

```{r}

d <- read_csv('distance_matrix_range_newline.csv') %>% column_to_rownames(var = '...1')
# make NAs into blanks
d[is.na(d)] <- " "
```

```{r}
#print table and save png
d %>% kbl(caption = "Uncorrected cytb Jukes-Cantor Differences   mean (min, max)", longtable = T, escape = F, align = 'c') %>% 
  kable_classic(#full_width = F, 
               html_font = "Cambria", 
               font_size = 10, 
               position = 'center') %>% 
 #column_spec(position = 'center')
  save_kable(file = "cytb_differences.png",zoom = 2)
```

# Print Matrix fewer species

This matrix doesn't have the morphologically and phylogenetically distant paradoxura and abscondita or maxi (close to elongata)

```{r}
d <- read_csv('distance_matrix_range_no_para_abs.csv') %>% column_to_rownames(var = '...1')
# make NAs into blanks
d[is.na(d)] <- " "
```

```{r}
#print table and save png
d %>% kbl(caption = "cytb Jukes-Cantor Differences   mean (min, max)") %>% 
  kable_classic(full_width = F, html_font = "Cambria", font_size = 10) %>%
  #as_image(width = 8, height = 5) %>% 
  save_kable(file = "cytb_differences_reduced.png",zoom = 2)
```


# Sumatra Only

```{r}
d <- read_csv('distance_matrix_range_Sumatra.csv') %>% column_to_rownames(var = '...1')
# make NAs into blanks
d[is.na(d)] <- " "
```

```{r}
#print table and save png
d %>% kbl(caption = "uncorrected cytb Jukes-Cantor Differences   mean (min, max)") %>% 
  kable_classic(full_width = F, html_font = "Cambria", font_size = 10) %>%
  #as_image(width = 8, height = 5) %>% 
  save_kable(file = "cytb_differences_Sumatra.png",zoom = 2)
```