---
title: "R Notebook"
output: html_notebook
---
Try some plotting!
Looking for differences in Sumatra shrews.

Below is a bunch of preliminary plotting using some confusing speciemns that I can't identify.
Now the new functional database is complete
```{r}
pacman::p_load(tidyverse, patchwork, kableExtra)
```
# Read Data

```{r}
db <- read_csv("Sumatra_DB.csv") %>% 
  mutate(Catalog = as.character(Catalog),
         HB = Total - Tail,
         RTL = Tail / HB) %>%
  rownames_to_column("id")

# Mass measurement for ordering species in plots
sp_ord <- db %>% 
  group_by(Sp2) %>% 
  filter(Sp2 != "unknown",
         Sp2 != "NA",
         !is.na(Mass)) %>% 
  summarise(.,mean = mean(Mass)) %>% 
  select(Sp2) %>% 
  as.list()
```
### Tables
```{r}
qtab <- function(metric) {
db %>% 
    #filter(Locality == "Talamau") %>%
    filter(Sp2 != "unknown") %>%
    mutate(Sp2 = recode(Sp2, "utongensis" = "neglecta")) %>%
    mutate(RSW = Condyloincisive / Brain_Breadth) %>%
    mutate(RRW = Rostral_Length / Rostral_Width) %>%
    mutate(IL = Upper_Tooth - P4_M3,
           IC = Interorbit_Width / Condyloincisive,
           SMW = Interorbit_Width / M2_M2_Labial) %>%
   #filter(Sp2 %in% c("dewi", "hutanis", "lepidura"), !is.na({{metric}})) %>%
   filter(Sp2 %in% c("barapi", "neglecta", "balingka"), !is.na({{metric}})) %>%
  select(Sp2, {{metric}}) %>% 
  group_by(Sp2) %>% 
  summarise(min({{metric}}),
            max({{metric}}),
            mean({{metric}}),
            sd({{metric}})) %>% 
    mutate_if(is.numeric, ~round(., 2))
    #add_column(metric = print({{metric}})) #%>% 
  #select(order(colnames(.))) %>% 
  #arrange(Condyloincisive) %>% 
   #kbl(digits = 2) %>%
  #kable_classic(full_width = F, html_font = "Cambria")
}
```
# Quick Table function
```{r}
qtab(HF) 
```

```{r}
qtab(Interorbit_Width)
```

# Skull Proportions
##Box plots
```{r}
hull_clade <- db %>% 
  filter(Sp2 != "unknown",
         !is.na(Brain_Breadth) & 
           !is.na(HB) &
           !is.na(Condyloincisive)) %>% 
  mutate(r_sk = Condyloincisive / Brain_Breadth) %>% 
  group_by(Sp2) %>% 
  slice(chull(HB, r_sk))

 db %>% 
  filter(Sp2 != "unknown",
         !is.na(Brain_Breadth) & 
           !is.na(Condyloincisive) &
           !is.na(HB)
         ) %>%
  mutate(r_sk = Condyloincisive / Brain_Breadth) %>% 
  mutate(Sp2 = recode(Sp2, "utongensis" = "neglecta")) %>% 
  ggplot(aes(x = HB,  y = r_sk, color = Sp2, fill = Sp2, shape = Sp2)) +
  geom_point(size = 3, stroke = 1)  +
  geom_polygon(data = hull_clade, alpha = 0.5) +
  scale_shape_manual(values = rep(1:10, 5)) +
  scale_fill_brewer(palette = "Dark2") +
  scale_color_brewer(palette = "Dark2") +
  theme_bw()
```

RObusticity


```{r}
 db %>% 
  filter(Sp2 != "unknown",
         !is.na(Brain_Breadth) & !is.na(Condyloincisive)
         ) %>%
  mutate(r_sk = Condyloincisive / Brain_Breadth) %>% 
  mutate(Sp2 = recode(Sp2, "utongensis" = "neglecta")) %>% 
  ggplot(aes(y = r_sk, x=reorder(Sp2,Mass, na.rm = TRUE),color = Sp2, fill = Sp2)) +
  geom_boxplot(alpha = 0.5) +
  scale_fill_brewer(palette = "Dark2") +
  scale_color_brewer(palette = "Dark2") +
  theme_bw() + 
  scale_x_discrete(guide = guide_axis(n.dodge = 2))
```


#### Rostrum Length vs Upper_Tooth
```{r}
hull_clade <- db %>% 
  filter(Sp2 != "unknown",
         !is.na(Rostral_Length) & 
           !is.na(Upper_Tooth)) %>% 
  group_by(Sp2) %>% 
  slice(chull(Rostral_Length, Upper_Tooth))

 db %>% 
  filter(Sp2 != "unknown",
         !is.na(Rostral_Length) & 
           !is.na(Upper_Tooth)) %>% 
  mutate(Sp2 = recode(Sp2, "utongensis" = "neglecta")) %>% 
  ggplot(aes(x = Rostral_Length,  y = Upper_Tooth, color = Sp2, fill = Sp2, shape = Sp2)) +
  geom_point(size = 3, stroke = 1)  +
  geom_polygon(data = hull_clade, alpha = 0.5) +
  scale_shape_manual(values = rep(1:10, 5)) +
  scale_fill_brewer(palette = "Dark2") +
  scale_color_brewer(palette = "Dark2") +
  theme_bw()
```

```{r}
db %>% 
  filter(Sp2 %in% c("barapi","balingka")) %>% 
  select(Sp2, Upper_Tooth, Rostral_Length, Interorbit_Width, M2_M2_Labial) %>% 
  arrange(Sp2, M2_M2_Labial) %>% 
   kbl(digits = 2) %>%
  kable_classic(full_width = F, html_font = "Cambria")
```

#### Rostrum width
```{r}
hull_clade <- db %>% 
  filter(Sp2 != "unknown",
         !is.na(Rostral_Length) & 
           !is.na(Rostral_Width)) %>% 
  group_by(Sp2) %>% 
  slice(chull(Rostral_Length, Rostral_Width))

 db %>% 
  filter(Sp2 != "unknown",
         !is.na(Rostral_Length) & 
           !is.na(Rostral_Width)) %>% 
  mutate(Sp2 = recode(Sp2, "utongensis" = "neglecta")) %>% 
  ggplot(aes(x = Rostral_Length,  y = Rostral_Width, color = Sp2, fill = Sp2, shape = Sp2)) +
  geom_point(size = 3, stroke = 1)  +
  geom_polygon(data = hull_clade, alpha = 0.5) +
  scale_shape_manual(values = rep(1:10, 5)) +
  scale_fill_brewer(palette = "Dark2") +
  scale_color_brewer(palette = "Dark2") +
  theme_bw()
 
 
  db %>% 
  filter(Sp2 != "unknown",
         !is.na(Rostral_Length) & !is.na(Rostral_Width)
         ) %>%
  mutate(Rostral_Shape = Rostral_Length / Rostral_Width) %>% 
  mutate(Sp2 = recode(Sp2, "utongensis" = "neglecta")) %>% 
  ggplot(aes(x=reorder(Sp2,Mass, na.rm = TRUE), y = Rostral_Shape, color = Sp2, fill = Sp2)) +
  geom_boxplot(alpha = 0.5, outlier.shape = NA) +
  geom_jitter(shape=16, position=position_jitter(0.2), alpha = 0.7)  +
  #geom_polygon(data = hull_clade, alpha = 0.5) +
  #scale_shape_manual(values = rep(1:10, 5)) +
  scale_fill_brewer(palette = "Dark2") +
  scale_color_brewer(palette = "Dark2") +
  theme_bw() + 
  scale_x_discrete(guide = guide_axis(n.dodge = 2))
```


# Hutanis Problem 1
### What is hutanis? What is beccarii?

PLots of Ruedi external limits vs. our reality
```{r}
db %>% 
  filter(Island == "Sumatra",
         Species != "unknown",
         Species %in% c("beccarii","hutanis", "lepidura"),
         #Locality == "Singgalang"
         Locality %in% c('Singgalang', 'Talamau', 'Tujuh') 
         ) %>%
  mutate(Species = recode(Species, "utongensis" = "neglecta")) %>% 
  ggplot(aes(y = Tail, x = HB, color = Species, shape = Locality)) +
  geom_point(size = 3, alpha = 0.85) + 
  geom_segment(aes(x = 73, y = 40, xend = 73, yend = 61), color = 'black', linetype = "dashed", size = 0.3) +
  #geom_segment(aes(x = 4, y = 15, xend = 4, yend = 27)) + 
  #geom_segment(aes(x = 4, y = 15, xend = 4, yend = 27)) +
  geom_vline(xintercept = 86.5, linetype = "dashed") +
  geom_hline(yintercept = 61, linetype = "dashed") + 
  scale_color_brewer(palette = 'Dark2', 
                     name = "mtDNA 'Species'") +
  annotate(geom="text", x=67, y=58, label="beccarii") +
  annotate(geom="text", x=81, y=58, label="hutanis") +
  annotate(geom="text", x=100, y=65, label="lepidura") +
  coord_cartesian(ylim = c(45, 80)) +
  labs(title = "Ruedi Measurement Scheme", subtitle = "Colors represent mtDNA clusters. \nLines represent the morphological boundaries described by Ruedi. \nnDNA provides no resolution, all are jumbled together and mixed with vosmaeri") +
  theme_classic()
ggsave("Ruedi_Morpho_CytB_plot.pdf", height = 5, width = 7)
```


```{r}
db %>% 
  filter(Island == "Sumatra",
         Species != "unknown",
         Species %in% c("beccarii","hutanis", "lepidura"),
         #Locality == "Singgalang"
         ) %>%
  mutate(Species = recode(Species, "utongensis" = "neglecta")) %>% 
  ggplot(aes(y = HF, x = HB, color = Species)) +
  geom_point(size = 3) + 
  geom_vline(xintercept = 73, linetype = "dashed") +
  geom_vline(xintercept = 86.5, linetype = "dashed") +
  geom_hline(yintercept = 16, linetype = "dashed") +
  geom_hline(yintercept = 13.1, linetype = "dashed") +
  annotate(geom="text", x=62, y=12.5, label="beccarii") +
  annotate(geom="text", x=80, y=15, label="hutanis") +
  annotate(geom="text", x=100, y=17, label="lepidura") +
  labs(title = "Ruedi Measurement Scheme", subtitle = "Colors represent mtDNA clusters.\nnDNA provides no resolution, all jumbled and mixed with vosmaeri") +
  theme_bw()
#ggsave("bec-hut-lep-HB-HF.pdf")
```
As can be seen in both of these, the measurements laid out by Ruedi don't make sense.

We can rename the species based on Ruedi, then see how well things fit.
```{r}
db %>% 
  mutate(Species = Sp2) %>% 
  mutate(Species = case_when(Species == "hutanis"  & HB   < 73 ~ "beccarii",
                             Species == "beccarii" & HB   > 73 ~ "hutanis",
                             Species == "hutanis"  & Tail > 62 ~ "lepidura",
                             Species == "unknown"  ~ "unknown",
                             TRUE ~ Species)) %>% 
  filter(Island == "Sumatra",
         #Species != "unknown",
         Species %in% c("beccarii","hutanis", "lepidura", "unknown"),
         #Locality == "Singgalang"
         ) %>%
  mutate(Species = recode(Species, "utongensis" = "neglecta")) %>% 
  ggplot(aes(y = Tail, x = HB, color = Species)) +
  geom_point(size = 3) + 
  geom_vline(xintercept = 73, linetype = "dashed") +
  geom_vline(xintercept = 86.5, linetype = "dashed") +
  geom_hline(yintercept = 61, linetype = "dashed") + 
  annotate(geom="text", x=62, y=58, label="beccarii") +
  annotate(geom="text", x=80, y=58, label="hutanis") +
  annotate(geom="text", x=100, y=65, label="lepidura") +
  labs(title = "Ruedi Measurement Scheme", subtitle = "Colors represent morphological clusters") +
  theme_bw()
```

```{r}
db %>% 
  mutate(Species = Species) %>% 
  mutate(Species = case_when(Species == "hutanis"  & HB   < 73 ~ "beccarii",
                             Species == "beccarii" & HB   > 73 ~ "hutanis",
                             Species == "hutanis"  & Tail > 62 ~ "lepidura",
                             TRUE ~ Species)) %>% 
  filter(Island == "Sumatra",
         Species != "unknown",
         Species %in% c("beccarii","hutanis", "lepidura"),
         #Locality == "Singgalang"
         ) %>%
  mutate(Species = recode(Species, "utongensis" = "neglecta")) %>% 
  ggplot(aes(y = HF, x = HB, color = Species)) +
  geom_point(size = 3) + 
  geom_vline(xintercept = 73, linetype = "dashed") +
  geom_vline(xintercept = 86.5, linetype = "dashed") +
  geom_hline(yintercept = 16, linetype = "dashed") +
  geom_hline(yintercept = 13.1, linetype = "dashed") +
  annotate(geom="text", x=62, y=12.5, label="beccarii") +
  annotate(geom="text", x=80, y=15, label="hutanis") +
  annotate(geom="text", x=100, y=17, label="lepidura") +
  labs(title = "Ruedi Measurement Scheme", subtitle = "Colors represent morpho clusters") +
  theme_bw()
```

```{r}
#db %>% filter(Island == "Sumatra") %>% 
#  write_csv("Sumatra_DB.csv")
```







```{r}
db %>% 
  mutate(Species = recode(Species, "utongensis" = "neglecta")) %>%
  filter(#Island == "Sumatra",
         #Species != "unknown"
         Species %in% c("barapi", "balingka", "neglecta")
         ) %>% 
  ggplot(aes(y = Condyloincisive, x = Species, color = Species)) +
  geom_violin() +
    geom_point() + 
  theme_bw()
```

```{r}
#options(digits = 2)
db %>% 
  filter(Island == "Sumatra",
         Sp2 != "unknown"#,
         #Locality == "Singgalang"
         ) %>% 
  mutate(Sp2 = recode(Sp2, "utongensis" = "neglecta")) %>%
  group_by(Sp2) %>% 
  summarise(RTL = mean(Tail/HB, na.rm = TRUE),
            sd = sd(Tail/HB, na.rm = TRUE)) %>% 
  as_tibble()
```
```{r}
 db %>% 
  filter(Sp2 != "unknown"#,
         #!is.na(Brain_Breadth) & !is.na(Condyloincisive)
         ) %>%
  mutate(RTL = Tail/HB) %>% 
  mutate(Sp2 = recode(Sp2, "utongensis" = "neglecta")) %>% 
  ggplot(aes(y = RTL, x=reorder(Sp2,Mass, na.rm = TRUE),color = Sp2, fill = Sp2)) +
  geom_boxplot(alpha = 0.5) +
  #geom_point(size = 3, stroke = 1)  +
  #geom_polygon(data = hull_clade, alpha = 0.5) +
  #scale_shape_manual(values = rep(1:10, 5)) +
  scale_fill_brewer(palette = "Dark2") +
  scale_color_brewer(palette = "Dark2") +
  theme_bw() + 
  scale_x_discrete(guide = guide_axis(n.dodge = 2))
```


# Generate mean and sd csv files
```{r}
#options(digits = 2)
db %>% 
  filter(
         Island == "Sumatra",
         Species != "unknown",
         #Locality == "Singgalang"
         ) %>% 
  mutate(Species = recode(Sp2, "utongensis" = "neglecta"),
         RTL = Tail / HB) %>%
  group_by(Sp2) %>% 
  summarise_if(is.numeric,  list(mean = mean, sd = sd), na.rm = TRUE) %>% 
  select(order(colnames(.))) %>%
  select(Sp2, everything()) %>% 
  write_csv("means_sds.csv")

db %>% 
  filter(
         Island == "Sumatra",
         Species != "unknown",
         Locality == "Talamau"
         ) %>% 
  mutate(Sp2 = recode(Sp2, "utongensis" = "neglecta"),
         RTL = Tail / HB) %>%
  group_by(Sp2) %>% 
  summarise_if(is.numeric,  list(mean = mean, sd = sd), na.rm = TRUE) %>% 
  select(order(colnames(.))) %>% 
  write_csv("means_sds_Talamau.csv")

db %>% 
  filter(
         Island == "Sumatra",
         Species != "unknown",
         Locality == "Singgalang"
         ) %>% 
  mutate(Sp2 = recode(Sp2, "utongensis" = "neglecta"),
         RTL = Tail / HB) %>%
  group_by(Sp2) %>% 
  summarise_if(is.numeric,  list(mean = mean, sd = sd), na.rm = TRUE) %>% 
  select(order(colnames(.))) %>% 
  write_csv("means_sds_Singgalang.csv")
```

# Body size and elevation

```{r}
db %>% 
  filter(
         Island == "Sumatra",
         Species  %in% c("beccarii", "hutanis", "lepidura"),
         Locality == "Talamau"
         ) %>% 
  ggplot(aes(y = Elevation, x = Mass)) + 
  geom_point(size = 3, stroke = 1) +
    theme_bw() +
    theme(legend.text = element_text(face = "italic")) +
   labs(
    y = "Elevation",
    x = "Mass"
  ) + ggtitle('Talamau hutanis/beccarii/lepidura, mass x elevation')

```

