---
title: "R Notebook"
output: html_notebook
---

# Plotting For Manuscript

The preliminary plots are in `Exploratory_Plotting.Rmd`. Based on these results, there are 4 plots that I really want to include: Linear PC1&2, Dorsal PC1&2, Ventral PC1&2, and the Condyloincisive x Tail plot. 
```{r}
pacman::p_load(tidyverse, patchwork, brms, bayesplot, tidybayes)
scale2 <- function(x, na.rm = FALSE) (x - mean(x, na.rm = na.rm)) / sd(x, na.rm)
```
# Read Data

```{r}
db <- read_csv("Sumatra_DB.csv") %>% 
  mutate(Catalog = as.character(Catalog),
         HB = Total - Tail,
         RTL = Tail / HB) %>%
  mutate(Species = recode(Species, "utongensis" = "neglecta")) %>% 
  rownames_to_column("id")

# Make a spreadsheet for the Sup Table

db %>% 
  select(Museum, 
         Catalog, 
         Species, 
         Condyloincisive:HB, 
         Sex, 
         Elevation, 
         Locality, 
         LPC1:LPC12, 
         DPCA1:VCsize) %>%
  rename_with(~str_replace(., 'LPC', 'Linear_PC')) %>% 
  rename_with(~str_replace(., 'DPCA', 'Dorsal_PC')) %>% 
  rename_with(~str_replace(., 'VPCA', 'Ventral_PC')) %>%
  rename_with(~str_replace(., 'DCsize', 'Dorsal_Centroid_Size')) %>%
  rename_with(~str_replace(., 'VCsize', 'Ventral_Centroid_Size')) %>%
  rename_with(~str_replace(., 'HF', 'Hind_Foot')) %>%
  rename_with(~str_replace(., 'HB', 'Head_Body')) %>%
  write_csv("Appendix_2_Supporting_Table_Specimen_Data.csv")
# Mass measurement for ordering species in plots
sp_ord <- db %>% 
  group_by(Species) %>% 
  filter(Species != "unknown",
         Species != "NA",
         !is.na(Mass)) %>% 
  summarise(.,mean = mean(Mass)) %>% 
  select(Species) %>% 
  as.list()
```

## Set Layers for repeated use
First time trying this
It's tricky, so I am just using the original full text plots
```{r}
lays = list(
  #geom_vline(xintercept = 0, linetype = 3, size = 0.25),
    #geom_hline(yintercept = 0, linetype = 3, size = 0.25),
    geom_point(size = 3, stroke = 1),
    scale_shape_manual(values = rep(1:10, 5), name = "Species"),
    scale_fill_brewer(palette = "Dark2", name = "Species"),
    scale_color_brewer(palette = "Dark2", name = "Species"),
    theme_bw(),
    theme(legend.text = element_text(face = "italic", size = 12))
  )
```


## Linear PC Plot

Plot PCA 1 & 2
```{r}
hull_clade <- db %>% 
  filter(Species != "unknown",
         !is.na(LPC1) & !is.na(LPC2)) %>% 
  group_by(Species) %>% 
  slice(chull(LPC1, LPC2))
  
p1 <- db %>% 
  filter(Species != "unknown",
         !is.na(LPC1) & !is.na(LPC2)
         ) %>%
  ggplot(aes(x = LPC1, y = LPC2, color = Species, fill = Species, shape = Species)) +
  geom_polygon(data = hull_clade, alpha = 0.4) +
  geom_point(size = 3, stroke = 1) +
    scale_shape_manual(values = rep(1:10, 5), name = "Species") +
    scale_fill_brewer(palette = "Dark2", name = "Species") +
    scale_color_brewer(palette = "Dark2", name = "Species") +
    theme_bw() +
    theme(legend.text = element_text(face = "italic", size = 12)) +
   labs(
    x = "Linear PC1 (91%)",
    y = "Linear PC2 (3%)"
  )
```

## Dorsal PC Plot

```{r}
hull_clade <- db %>% 
  filter(Species != "unknown",
         !is.na(DPCA1) & !is.na(DPCA2)) %>% 
  group_by(Species) %>% 
  slice(chull(DPCA1, DPCA2))
  
p2 <- db %>% 
  filter(Species != "unknown",
         !is.na(DPCA1) & !is.na(DPCA2)
         ) %>%
  ggplot(aes(x = DPCA1, y = DPCA2, color = Species, fill = Species, shape = Species)) +
  geom_polygon(data = hull_clade, alpha = 0.4) +
  geom_point(size = 3, stroke = 1) +
    scale_shape_manual(values = rep(1:10, 5), name = "Species") +
    scale_fill_brewer(palette = "Dark2", name = "Species") +
    scale_color_brewer(palette = "Dark2", name = "Species") +
    theme_bw() +
    theme(legend.text = element_text(face = "italic", size = 12)) +
   labs(
    x = "Dorsal PC1 (38%)",
    y = "Dorsal PC2 (14%)"
  )
```

## Ventral PC Plot

```{r}
hull_clade <- db %>% 
  filter(Species != "unknown",
         !is.na(VPCA1) & !is.na(VPCA2)) %>% 
  group_by(Species) %>% 
  slice(chull(VPCA1, VPCA2))
  
p3 <- db %>% 
  filter(Species != "unknown",
         !is.na(VPCA1) & !is.na(VPCA2)
         ) %>%
  ggplot(aes(x = VPCA1, y = VPCA2, color = Species, fill = Species, shape = Species)) +
  geom_polygon(data = hull_clade, alpha = 0.4) +
  geom_point(size = 3, stroke = 1) +
    scale_shape_manual(values = rep(1:10, 5), name = "Species") +
    scale_fill_brewer(palette = "Dark2", name = "Species") +
    scale_color_brewer(palette = "Dark2", name = "Species") +
    theme_bw() +
    theme(legend.text = element_text(face = "italic", size = 12)) +
   labs(
    x = "Ventral PC1 (35%)",
    y = "Ventral PC2 (11%)"
  )
```

## CI and Tail
Condyloincisive and Tail is a very nice bivariate plot. This does not need the 0-0 lines, so I am not using the `lays`

Geom_smooth doesn't go to the edge, so I am using abline

```{r}
dd <- db %>% 
  filter(Species != "unknown",
         !is.na(Condyloincisive) & !is.na(Tail),
         Species != "paradoxura"
         ) 
lm(Condyloincisive ~ Tail, dd)
```

```{r}
hull_clade <- db %>% 
  filter(Species != "unknown",
         !is.na(Condyloincisive) & !is.na(Tail)) %>% 
  group_by(Species) %>% 
  slice(chull(Condyloincisive, Tail))
  
p4 <-  db %>% 
  filter(Species != "unknown",
         !is.na(Condyloincisive) & !is.na(Tail)
         ) %>%
  ggplot(aes(y = Condyloincisive, x = Tail, color = Species, fill = Species, shape = Species)) + 
  geom_abline(slope =  0.2423 , 
              intercept = 6.7275, 
              color="black", 
              size=0.3, 
              linetype = 5) +
  geom_polygon(data = hull_clade, alpha = 0.4) +
  geom_point(size = 3, stroke = 1) +
    scale_shape_manual(values = rep(1:10, 5), name = "Species") +
    scale_fill_brewer(palette = "Dark2", name = "Species") +
    scale_color_brewer(palette = "Dark2", name = "Species") +
    theme_bw() +
    theme(legend.text = element_text(face = "italic", size = 12)) +
   labs(
    y = "Condyloincisive Length (mm)",
    x = "Tail Length (mm)"
  ) + xlim(33,78) #+
   #coord_cartesian(xlim = c(33, 78))
p4
```

## CI and Tail 2
Using Paradoxura, but not using the regression line. 
```{r}
dd <- db %>% 
  mutate(ltail = log(Tail)) 
```

```{r}
hull_clade <- dd %>% 
  filter(Species != "unknown",
         !is.na(Condyloincisive) & !is.na(ltail)) %>% 
  group_by(Species) %>% 
  slice(chull(Condyloincisive, ltail))
  
p4 <-  dd %>% 
  filter(Species != "unknown",
         !is.na(Condyloincisive) & !is.na(ltail)
         ) %>%
  ggplot(aes(y = Condyloincisive, x = ltail, color = Species, fill = Species, shape = Species)) + 
  #geom_abline(slope =  0.2423 , 
  #            intercept = 6.7275, 
  #            color="black", 
  #            size=0.3, 
  #            linetype = 5) +
  geom_polygon(data = hull_clade, alpha = 0.4) +
  geom_point(size = 3, stroke = 1) +
    scale_shape_manual(values = rep(1:10, 5), name = "Species") +
    scale_fill_brewer(palette = "Dark2", name = "Species") +
    scale_color_brewer(palette = "Dark2", name = "Species") +
    theme_bw() +
    theme(legend.text = element_text(face = "italic", size = 12)) +
   labs(
    y = "Condyloincisive Length (mm)",
    x = "log(Tail Length (mm))"
  ) #+ xlim(33,78) #+
   #coord_cartesian(xlim = c(33, 78))
p4
```
# Big Plot

Linear Measurements on top?

```{r}
( p1 | p4) / (p2 | p3) + plot_layout(guides = "collect") + plot_annotation(tag_levels = 'A') 
ggsave('Morpho_Figure.pdf', height = 9.2, width = 11)
```




# Skull Length vs Skull Width
```{r}

hull_clade <- db %>% 
  mutate(Rostral_Width = log(Rostral_Width)) %>% 
  filter(Species != "unknown",
         !is.na(Condyloincisive) & !is.na(Rostral_Width)) %>% 
  group_by(Species) %>% 
  slice(chull(Condyloincisive, Rostral_Width))
  
p4 <-  db %>% 
  mutate(Rostral_Width = log(Rostral_Width)) %>% 
  filter(Species != "unknown",
         !is.na(Condyloincisive) & !is.na(Rostral_Width)
         ) %>%
  ggplot(aes(y = Condyloincisive, x = Rostral_Width, color = Species, fill = Species, shape = Species)) + 
  geom_polygon(data = hull_clade, alpha = 0.4) +
  geom_point(size = 3, stroke = 1) +
    scale_shape_manual(values = rep(1:10, 5), name = "Species") +
    scale_fill_brewer(palette = "Dark2", name = "Species") +
    scale_color_brewer(palette = "Dark2", name = "Species") +
    theme_bw() +
    theme(legend.text = element_text(face = "italic", size = 12)) +
   labs(
    y = "Condyloincisive Length",
    x = "Rostral_Width Length"
  ) 

p4
```



# RTL or TAIL?

```{r}
hull_clade <- db %>% 
  filter(Species != "unknown",
         !is.na(Condyloincisive) & !is.na(RTL)) %>% 
  group_by(Species) %>% 
  slice(chull(Condyloincisive, RTL))
  
p5 <-  db %>% 
  filter(Species != "unknown",
         !is.na(Condyloincisive) & !is.na(RTL)
         ) %>%
  ggplot(aes(y = Condyloincisive, x = RTL, color = Species, fill = Species, shape = Species)) + 
  geom_polygon(data = hull_clade, alpha = 0.4) +
  geom_point(size = 3, stroke = 1) +
    scale_shape_manual(values = rep(1:10, 5), name = "Species") +
    scale_fill_brewer(palette = "Dark2", name = "Species") +
    scale_color_brewer(palette = "Dark2", name = "Species") +
    theme_bw() +
    theme(legend.text = element_text(face = "italic", size = 12)) +
   labs(
    y = "Condyloincisive Length",
    x = "R Tail Length"
  ) + xlim(0.5, 1)
```
 

```{r}
p4 / p5 + plot_layout(guides = "collect") 
```


## m2m2 labial
```{r}
hull_clade <- db %>% 
  filter(Species != "unknown",
         !is.na(Condyloincisive) & !is.na(M2_M2_Labial)) %>% 
  mutate(Condyloincisive = scale2(Condyloincisive),
         M2_M2_Labial = scale2(M2_M2_Labial)) %>% 
  group_by(Species) %>% 
  slice(chull(Condyloincisive, M2_M2_Labial))
  
p5 <-  db %>% 
  filter(Species != "unknown",
         !is.na(Condyloincisive) & !is.na(M2_M2_Labial)
         ) %>%
    mutate(Condyloincisive = scale2(Condyloincisive),
         M2_M2_Labial = scale2(M2_M2_Labial)) %>% 
  ggplot(aes(y = Condyloincisive, x = M2_M2_Labial, color = Species, fill = Species, shape = Species)) + 
  geom_polygon(data = hull_clade, alpha = 0.4) +
  geom_point(size = 3, stroke = 1) +
    scale_shape_manual(values = rep(1:10, 5), name = "Species") +
    scale_fill_brewer(palette = "Dark2", name = "Species") +
    scale_color_brewer(palette = "Dark2", name = "Species") +
    theme_bw() +
  geom_abline(slope = 1, intercept = 0) +
    theme(legend.text = element_text(face = "italic", size = 12)) +
   labs(
    y = "Condyloincisive Length",
    x = "M2_M2_Labial"
  ) #+ xlim(0.5, 1)
p5
```
