---
title: "Sumatra Mapping"
author: "Jon Nations"
date: "`r Sys.Date()`"
output: html_document
---

Mapping the Localities from Recent Sumatra Surveys

I will start by using the Coepernicus DEM data from the [Eurostat DEM](https://ec.europa.eu/eurostat/web/gisco/geodata/reference-data/elevation/copernicus-dem/elevation) website. Turns out this stuff Rules!

TODO: - Merge IDN and Malaysia and Singapore Rasters and Shape Data so that Peninsula shows up on the large map. This is the 

Packages
```{r}
pacman::p_load(tidyverse, sf, raster, spData, spDataLarge, tmap, leaflet, grDevices, rgdal, gdalUtils, rgeos)
```




## Reading In Raster and Shape Data 

```{r}
#Get basic elevation raster
idn_alt <- getData('alt', country='IDN', mask=TRUE)
mys_alt <- getData('alt', country='MYS', mask=TRUE)
tha_alt <- getData('alt', country='THA', mask=TRUE)
#sgp_alt <- getData('alt', country='SGP', mask=TRUE)
idn_alt <- raster::merge(idn_alt, mys_alt, tha_alt)

#Get shape data
idn_gadm <- getData('GADM', country='IDN', level = 1)
mys_gadm <- getData('GADM', country='MYS', level = 1)
tha_gadm <- getData('GADM', country='THA', level = 1)
#sgp_gadm <- getData('GADM', country='SGP', level = 1)
idn_gadm <- union(idn_gadm, mys_gadm)
idn_gadm <- union(idn_gadm, tha_gadm)
```


bounding box and projection
```{r}
idn_region = st_bbox(c(xmin = 93, 
                       xmax = 107, 
                       ymax = 7, 
                       ymin = -8),
                     crs = st_crs(idn_alt)) %>% 
   st_as_sfc()
```

# Big Sumatra Map

## Large Map

Custom palette
```{r}
palls <- terrain.colors(100)
pallsbw = gray.colors(100, start = 0.9, end =0)
#the best
specialbw <- c("grey90", "grey70", "grey60", "grey30", "grey15", "grey5")
```

Trying with data downloaded from  `raster` package - maybe it's sufficient? Sumatra is much bigger than Sulawesi
```{r}
idn_alt2 <- abs(idn_alt)

map1 <- tm_shape(idn_alt2, bbox = idn_region) + 
  tm_raster(title = "Elevation (m)", 
            alpha = 0.9, 
            #this makes a continuous range for the color palette
           # style = "cont", 
            midpoint = NA, 
            palette = specialbw, 
            #palette = pallsbw,
            breaks = c(0, 500, 1000, 1500, 2000, 2500, 4651)) +
            #labels = c ("","700m","","2100m","","3500m")) + 
  tm_shape(idn_gadm, 
           bbox = idn_region) + 
  tm_polygons(border.col = "black", 
              alpha = 0) + 
  tm_scale_bar(position = c("center", "bottom"), 
               text.size = 1, 
               color.dark = "black", 
               color.light = "grey50", 
               lwd = 1) + 
  tm_layout(legend.title.size = 1.5,
            legend.text.size = 0.8,
            legend.position = c("left","bottom"),
            legend.bg.color = "white",
            legend.bg.alpha = 1) + 
  tm_graticules(lines = FALSE)

#map1

tmap_save(map1, file = "big_sumatra_map.pdf")
```

### Different number of breaks
```{r}
idn_alt2 <- abs(idn_alt)
specialbw2 <- c("white", "grey92", "grey50", "grey25", "grey12", "grey5")
#specialbw <- c("grey95", "grey50","grey250", "grey12", "grey5")
map1 <- tm_shape(idn_alt2, bbox = idn_region) + 
  tm_raster(title = "Elevation (m)", 
            alpha = 0.9, 
            #this makes a continuous range for the color palette
           # style = "cont", 
            midpoint = NA, 
            palette = specialbw2, 
            #palette = pallsbw,
            breaks = c(0, 2, 700, 1400, 2100, 2700, 4651)) +
            #labels = c ("","700m","","2100m","","3500m")) + 
  tm_shape(idn_gadm, 
           bbox = idn_region) + 
  tm_polygons(border.col = "black", 
              alpha = 0, lwd = 0.1,
              border.alpha = 0.6) + 
  tm_scale_bar(position = c("center", "bottom"), 
               text.size = 1, 
               color.dark = "black", 
               color.light = "grey50", 
               lwd = 1) + 
  tm_layout(legend.title.size = 1.5,
            legend.text.size = 0.8,
            legend.position = c("left","bottom"),
            legend.bg.color = "white",
            legend.bg.alpha = 1) + 
  tm_graticules(lines = FALSE)

#map1

tmap_save(map1, file = "big_sumatra_map2.pdf")
```


# WEST SUMATRA ZOOM 
#### Merged DEM

Just to try a smaller area to see the value of the Rasters.

Looks like i need better rasters. LEt's see what I can download.

## Merge Coepernicus rasters 

#### Only Needed to be Done Once! 

Here I am merging the tiff rasters. I downloaded the ones I need from the website above. I am trying the solution answer from [this post](https://stackoverflow.com/questions/15876591/merging-multiple-rasters-in-r)

`#``{r eval = FALSE}
r1 <- raster("10_DEM_y0x100.tif")
r2 <- raster("10_DEM_y0x90.tif")
r3 <- raster("10_DEM_y-10x100.tif")
r4 <- raster("10_DEM_y-10x90.tif")

x <- list(r1, r2, r3, r4)
names(x) <- c("x", "y")
x$filename <- 'Sumatra_Raster.tif'
x$overwrite <- TRUE
m <- do.call(merge, x)
#```



So for some reason the bbox isn't working.

I think I need to crop the raster SEPERATELY, then plot, not using bounding box, but I don't know why. 


```{r}

x <- raster("Sumatra_Raster.tif")
st_crs(x)
```
This is looking good!
```{r}
ws_region = st_bbox(c(xmin = 99.5, 
                       xmax = 101, 
                       ymax = 0.5, 
                       ymin = -1))


specialbw <- c("white", "grey94", "grey80", "grey72", "grey64", "grey58", "grey50", "grey40", "grey32", "grey25", "grey15", "grey5")

DEM = abs(crop(x = x, y = ws_region))

map2 <- tm_shape(DEM) + 
  tm_raster(title = "Elevation (m)", 
            alpha = 0.9, 
            #this makes a continuous range for the color palette
            #style = "cont", 
            midpoint = NA, 
            palette = specialbw, 
            #palette = pallsbw,
            breaks = c(0, 1, 500, 750, 1000, 1250, 1500, 1750, 2000, 2250, 2500, 2750, 3760),
            labels = c ("","0m", "", "", "1000m", "", "", "2000m", "", "", "3000m","")) + 
  tm_shape(idn_gadm, 
           bbox = idn_region) +
  tm_polygons(border.col = "black", 
              alpha = 0, 
              lwd = 0.1,
              border.alpha = 0.6) + 
  tm_scale_bar(position = c("left", "bottom"), 
               text.size = .8, 
               color.dark = "black", 
               color.light = "grey50", 
               lwd = 1.5) + 
  tm_layout(legend.title.size = 1,
            legend.text.size = 0.7,
            legend.position = c("left","center"),
            legend.bg.color = "white",
            legend.bg.alpha = 0) + 
  tm_graticules(n.x = 4,
                lines = FALSE) 


tmap_save(map2, file = "west_sumatra_map.pdf")
```

Try a sparser sumatra map
```{r}
ws_region = st_bbox(c(xmin = 99.5, 
                       xmax = 101, 
                       ymax = 0.5, 
                       ymin = -1))


specialbw2 <- c("white", "grey98", "grey50", "grey15")

DEM = abs(crop(x = x, y = ws_region))

map2 <- tm_shape(DEM) + 
  tm_raster(title = "Elevation (m)", 
            alpha = 0.9, 
            #this makes a continuous range for the color palette
            #style = "cont", 
            midpoint = NA, 
            palette = specialbw2, 
            #palette = pallsbw,
            breaks = c(0, 1,  1000,  2000, 3760),
            labels = c ("", "0m", "1000m", "2000m")) + 
  tm_shape(idn_gadm, 
           bbox = idn_region) +
  tm_polygons(border.col = "black", 
              alpha = 0, 
              lwd = 0.1,
              border.alpha = 0.6) + 
  tm_scale_bar(position = c("left", "bottom"), 
               text.size = .8, 
               color.dark = "black", 
               color.light = "grey50", 
               lwd = 1.5) + 
  tm_layout(legend.title.size = 1,
            legend.text.size = 0.7,
            legend.position = c("left","center"),
            legend.bg.color = "white",
            legend.bg.alpha = 0) + 
  tm_graticules(n.x = 4,
                lines = FALSE) 


tmap_save(map2, file = "west_sumatra_map_2.pdf")
```
And a bigger map!
```{r}
ws_region = st_bbox(c(xmin = 99, 
                       xmax = 101.5, 
                       ymax = 1, 
                       ymin = -1.5))


specialbw2 <- c("white", "grey98", "grey50", "grey15")

DEM = abs(crop(x = x, y = ws_region))

map2 <- tm_shape(DEM) + 
  tm_raster(title = "Elevation (m)", 
            alpha = 0.9, 
            #this makes a continuous range for the color palette
            #style = "cont", 
            midpoint = NA, 
            palette = specialbw2, 
            #palette = pallsbw,
            breaks = c(0, 1,  1000,  2000, 3760),
            labels = c ("", "0m", "1000m", "2000m")) + 
  tm_shape(idn_gadm, 
           bbox = idn_region) +
  tm_polygons(border.col = "black", 
              alpha = 0, 
              lwd = 0.7,
              border.alpha = 0.6) + 
  tm_scale_bar(position = c("left", "bottom"), 
               text.size = .8, 
               color.dark = "black", 
               color.light = "grey50", 
               lwd = 1.5) + 
  tm_layout(legend.title.size = 1,
            legend.text.size = 0.7,
            legend.position = c("left","center"),
            legend.bg.color = "white",
            legend.bg.alpha = 0) + 
  tm_graticules(n.x = 4,
                lines = FALSE) 


tmap_save(map2, file = "west_sumatra_map_3.pdf")
```

And a bigger map!-1.6918650480214894, 101.41197584482654
```{r}
ws_region = st_bbox(c(xmin = 99, 
                       xmax = 102, 
                       ymax = 1.25, 
                       ymin = -2.75))


specialbw2 <- c("white", "grey94", "grey50", "grey15")

DEM = abs(crop(x = x, y = ws_region))

map4 <- tm_shape(DEM) + 
  tm_raster(title = "Elevation (m)", 
            alpha = 0.9, 
            #this makes a continuous range for the color palette
            #style = "cont", 
            midpoint = NA, 
            palette = specialbw2, 
            #palette = pallsbw,
            breaks = c(0, 1,  1000,  2000, 3760),
            labels = c ("", "<1000m", "1000-2000m", ">2000m")) + 
  tm_shape(idn_gadm, 
           bbox = idn_region) +
  tm_polygons(border.col = "black", 
              alpha = 0, 
              lwd = 0.7,
              border.alpha = 0.6) + 
  tm_scale_bar(position = c("center", "bottom"), 
               text.size = .8, 
               color.dark = "black", 
               color.light = "grey50", 
               lwd = 1.5) + 
  tm_layout(legend.title.size = 1,
            legend.text.size = 0.7,
            legend.position = c("right","top"),
            legend.bg.color = "white",
            legend.bg.alpha = 0) + 
  tm_graticules(n.x = 4,
                lines = FALSE) 


tmap_save(map4, file = "west_sumatra_map_4.pdf")
```
inset map
```{r}
wr_region = st_bbox(c(xmin = 95, 
                       xmax = 130, 
                       ymax = 15, 
                       ymin = -15))
wrmap <- tm_shape(world, bbox = wr_region) + tm_polygons()

tmap_save(wrmap, file = "inset_map.pdf")
```



```{r}
library(stars)
rm(DEM)
tif = system.file("Sumatra_Raster.tif", proxy=)
(x = read_stars(tif))

#DEM <- raster("Sumatra_Raster.tif")
DEM <- read_stars(tif)
```

