---
title: "Tree Plotting"
output: html_notebook
---

# Plotting the trees for the sumatra description 

## Load Data and Packages
```{r}
pacman::p_load(treedataverse, tidyverse, ape, TDbook, ggbreak, patchwork, ggpubr)

```

```{r}
mcc <- read.beast('trees/species_MCC.tree')
```

hmm, long 'outgroup' branch, but deal with that later
What do I want to plot? Traits? Islands? Let's start with islands
Manually making a mcc_treedata.csv
```{r}
tdata <- read_csv('trees/mcc_treedata.csv')
```

#### Quick Plot 

```{r}
ggtree(mcc) + geom_tiplab()
```

## Tree with All Localities Points
Attach treedata and sample plot
```{r}

p <- ggtree(mcc) %<+% tdata 

p2 <- p + geom_tiplab(offset = 0.0006, size = 3.5) + 
    geom_tippoint(aes(x=x+0.0004, color = island, fill = island, shape = island), size = 3) + 
  geom_treescale(x = 0.01, y = 0, width=0.001, offset = 0.2, fontsize = 3) +
    #scale_size_continuous(range = c(3, 10)) 
     scale_fill_brewer(type = 'qual', palette = "Dark2") +
  scale_color_brewer(type = 'qual', palette = "Dark2") +
  xlim(0, 0.018)

legend <- as_ggplot(ggpubr::get_legend(p2))

p3 <- p2 + scale_x_break(c(0.001, 0.005)) + hexpand(1) 
p4 <- p3 + theme(legend.position = c(0.5, 0.5),
                 legend.title = element_blank())
#p5 <- p4   +   inset_element(legend, 0, 0.5, 0.2 ,1) 
p4
ggsave('trees/mcc_tree1.pdf')
```

```{r}
legend
ggsave('trees/mcc_legend.pdf')
```

```{r}
p4 + theme(legend.position = 'none')
ggsave('trees/mcc_tree1_no_legend.pdf')
```


Ok, how to make this better?




geom_nodepoint(aes( label=posterior, subset = !is.na(as.numeric(posterior)) & as.numeric(posterior) > 0.9), size = 1.7, color ="grey30", alpha = 0.8)

```{r}
p4 + 
  geom_nodepoint(aes( label=posterior, subset = !is.na(as.numeric(posterior)) & as.numeric(posterior) > 0.9), size = 1.7, color ="black", alpha = 1) + 
  geom_nodepoint(aes( label=posterior, subset = !is.na(as.numeric(posterior)) & as.numeric(posterior) > 0.5 & as.numeric(posterior) < 0.9), size = 1.7, color ="grey40", alpha = 1, shape = 15) +
  geom_nodelab(aes(x=branch, label=round(posterior, 2), 
                   subset = !is.na(as.numeric(posterior)) & 
                     as.numeric(posterior) < .5),
               vjust=-.5, 
               #hjust = .7, 
               size = 1.5)
ggsave('trees/mcc_tree1_nodes.pdf', width = 6, height = 6.5)
```

# Tree Sumatra Nodepoint

```{r}
tdata3 <- tdata %>% #column_to_rownames("label") %>%
  mutate(island = fauxnaif::na_if_in(island, "Mainland", "Borneo", "Sulawesi", "Philippines", "Java"))

p <- ggtree(mcc) %<+% tdata3
p <- ggtree::flip(p, getMRCA(mcc@phylo, c("hutanis", "brunnea")), getMRCA(mcc@phylo, c("foetida", "negrina")))
p <- ggtree::flip(p, getMRCA(mcc@phylo, c("monticola", "umbra")), getMRCA(mcc@phylo, c("balingka", "dewi")))

p2 <- p + geom_tiplab(offset = 0.0002, size = 3.5) + 
    geom_tippoint(aes(x=x+0.0001, color = island, fill = island, shape = island), size = 3) + 
  geom_treescale(x = 0.01, y = 0, width=0.001, offset = 0.2, fontsize = 3) +
    #scale_size_continuous(range = c(3, 10)) 
     scale_fill_brewer(type = 'qual', palette = "Dark2") +
  scale_color_brewer(type = 'qual', palette = "Dark2") +
  xlim(0, 0.018)

legend <- as_ggplot(ggpubr::get_legend(p2))

p3 <- p2 + scale_x_break(c(0.001, 0.005)) + hexpand(1) 
p4 <- p3 + theme(legend.position = "none",
                 legend.title = element_blank())
#p5 <- p4   +   inset_element(legend, 0, 0.5, 0.2 ,1) 
p5 <- p4 + geom_nodepoint(aes( label=posterior, subset = !is.na(as.numeric(posterior)) & as.numeric(posterior) > 0.9), size = 1.7, color ="black", alpha = 1) + 
  geom_nodepoint(aes( label=posterior, subset = !is.na(as.numeric(posterior)) & as.numeric(posterior) > 0.5 & as.numeric(posterior) < 0.9), size = 1.7, color ="grey40", alpha = 1, shape = 15) +
  geom_nodelab(aes(x=branch, label=round(posterior, 2), 
                   subset = !is.na(as.numeric(posterior)) & 
                     as.numeric(posterior) < .5),
               vjust=.9, 
               hjust = -.95, 
               size = 1.75)
p5
ggsave('trees/mcc_tree1_nodes_sum.pdf', width = 5, height = 6.5)
```


## Tree Plot All Localities Heatmap
```{r}
tdata2 <- tdata %>% column_to_rownames("label")
p <- ggtree(mcc) 
p <- ggtree::flip(p, getMRCA(mcc@phylo, c("hutanis", "brunnea")), getMRCA(mcc@phylo, c("foetida", "negrina")))
p <- ggtree::flip(p, getMRCA(mcc@phylo, c("monticola", "umbra")), getMRCA(mcc@phylo, c("balingka", "dewi")))#+ rotate(p2, 36)
p2 <- p + geom_tiplab(offset = 0.0006, size = 3.5) + 
    #geom_tippoint(aes(x=x+0.0004, color = island, fill = island, shape = island), size = 3) + 
  geom_treescale(x = 0.01, y = 0, width=0.001, offset = 0.2, fontsize = 3) +
  xlim(0, 0.018)
#p2 <-   ggtree::rotate(p2, 36)

#legend <- as_ggplot(ggpubr::get_legend(p2))

p3 <- p2 + scale_x_break(c(0.001, 0.005)) + hexpand(1) 
p4 <- p3 + theme(legend.position = c(0.5, 0.5),
                 legend.title = element_blank())
#p5 <- p4   +   inset_element(legend, 0, 0.5, 0.2 ,1) 
p4


p5 <- gheatmap(p4, tdata2, offset=0, width=0.025, font.size=2, colnames=F, hjust=0) + 
     scale_fill_brewer(type = 'qual', palette = "Dark2") +
  scale_color_brewer(type = 'qual', palette = "Dark2") 
p6 <- p5 + geom_nodepoint(aes( label=posterior, subset = !is.na(as.numeric(posterior)) & as.numeric(posterior) > 0.9), size = 1.7, color ="black", alpha = 1) + 
  geom_nodepoint(aes( label=posterior, subset = !is.na(as.numeric(posterior)) & as.numeric(posterior) > 0.5 & as.numeric(posterior) < 0.9), size = 1.7, color ="grey40", alpha = 1, shape = 15) +
  geom_nodelab(aes(x=branch, label=round(posterior, 2), 
                   subset = !is.na(as.numeric(posterior)) & 
                     as.numeric(posterior) < .5),
               vjust=.9, 
               hjust = -.95, 
               size = 1.75)
p6

ggsave('trees/mcc_tree1_heatmap1.pdf', width = 6, height = 6.5)
```

## Tree Plot Sumatra Java Heatmap
```{r}
p <- ggtree(mcc) 
p <- ggtree::flip(p, getMRCA(mcc@phylo, c("hutanis", "brunnea")), getMRCA(mcc@phylo, c("foetida", "negrina")))
p <- ggtree::flip(p, getMRCA(mcc@phylo, c("monticola", "umbra")), getMRCA(mcc@phylo, c("balingka", "dewi")))#+ rotate(p2, 36)
p2 <- p + geom_tiplab(offset = 0.0006, size = 3.5) + 
    #geom_tippoint(aes(x=x+0.0004, color = island, fill = island, shape = island), size = 3) + 
  geom_treescale(x = 0.01, y = 0, width=0.001, offset = 0.2, fontsize = 3) +
  xlim(0, 0.018)
#p2 <-   ggtree::rotate(p2, 36)

#legend <- as_ggplot(ggpubr::get_legend(p2))

p3 <- p2 + scale_x_break(c(0.001, 0.005)) + hexpand(1) 
p4 <- p3 + theme(legend.position = c(0.5, 0.5),
                 legend.title = element_blank())
#p5 <- p4   +   inset_element(legend, 0, 0.5, 0.2 ,1) 
p4

tdata3 <- tdata %>% column_to_rownames("label") %>%
  mutate(island = fauxnaif::na_if_in(island, "Mainland", "Borneo", "Sulawesi", "Philippines"))

p5 <- gheatmap(p4, tdata3, offset=0, width=0.025, font.size=2, colnames=F, hjust=0) + 
   scale_fill_manual(breaks=c( "Java", "Sumatra"), 
        values=c("#414487", "#7AD151"), na.value = "white")
p6 <- p5 + geom_nodepoint(aes( label=posterior, subset = !is.na(as.numeric(posterior)) & as.numeric(posterior) > 0.9), size = 1.7, color ="black", alpha = 1) + 
  geom_nodepoint(aes( label=posterior, subset = !is.na(as.numeric(posterior)) & as.numeric(posterior) > 0.5 & as.numeric(posterior) < 0.9), size = 1.7, color ="grey40", alpha = 1, shape = 15) +
  geom_nodelab(aes(x=branch, label=round(posterior, 2), 
                   subset = !is.na(as.numeric(posterior)) & 
                     as.numeric(posterior) < .5),
               vjust=.9, 
               hjust = -.95, 
               size = 1.75)
p6
ggsave('trees/mcc_tree1_heatmap2.pdf', width = 6, height = 6.5)
```

## Tree Plot Sumatra Heatmap
```{r}
p <- ggtree(mcc) 
p <- ggtree::flip(p, getMRCA(mcc@phylo, c("hutanis", "brunnea")), getMRCA(mcc@phylo, c("foetida", "negrina")))
p <- ggtree::flip(p, getMRCA(mcc@phylo, c("monticola", "umbra")), getMRCA(mcc@phylo, c("balingka", "dewi")))
p2 <- p + geom_tiplab(offset = 0.0003, size = 3.5) + 
    #geom_tippoint(aes(x=x+0.0004, color = island, fill = island, shape = island), size = 3) + 
  geom_treescale(x = 0.01, y = 0, width=0.001, offset = 0.2, fontsize = 3) +
  xlim(0, 0.018)
#p2 <-   ggtree::rotate(p2, 36)

#legend <- as_ggplot(ggpubr::get_legend(p2))

p3 <- p2 + scale_x_break(c(0.001, 0.005)) + hexpand(1) 
p4 <- p3 + theme(legend.position = c(0.5, 0.5),
                 legend.title = element_blank())
tdata3 <- tdata %>% column_to_rownames("label") %>%
  mutate(island = fauxnaif::na_if_in(island, "Mainland", "Borneo", "Sulawesi", "Philippines", "Java"))

p5 <- gheatmap(p4, tdata3, offset=-.000175, width=0.02, font.size=2, colnames=F, hjust=0) + 
   scale_fill_manual(breaks=c( "Sumatra"), 
        values=c("#7AD151"), na.value = "white") + theme(legend.position = 'none')
p6 <- p5 + geom_nodepoint(aes( label=posterior, subset = !is.na(as.numeric(posterior)) & as.numeric(posterior) > 0.9), size = 1.7, color ="black", alpha = 1) + 
  geom_nodepoint(aes( label=posterior, subset = !is.na(as.numeric(posterior)) & as.numeric(posterior) > 0.7 & as.numeric(posterior) < 0.9), size = 1.7, color ="grey40", alpha = 1, shape = 15) +
  geom_nodelab(aes(x=branch, label=round(posterior, 2), 
                   subset = !is.na(as.numeric(posterior)) & 
                     as.numeric(posterior) < .7),
               vjust=.9, 
               hjust = -.95, 
               size = 1.75)
p6
ggsave('trees/mcc_tree1_heatmap3.pdf', width = 5, height = 6.5)
```

